---
title: "Machine Learning Project"
author: "MLR"
date: "23. september 2016"
output: html_document
---

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform correctly and incorrectly in 5 different ways. 

## Preparing and investigating the data
```{r}
library(ggplot2); library(caret)

# Reading files
orgtrain <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
orgtest <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

# creating new dataframe with only the accelerometer data 
mtottrain <- orgtrain[,c(40,41,42,    #belt
                    63,64,65,    #arm
                    116,117,118, #dumbbell
                    154,155,156, #forearm
                    160)]        #classe

# Investigating the dataset
head(mtottrain)
str(mtottrain)
dim(mtottrain)
summary(mtottrain)
summary(is.na(mtottrain))

# Plotting
featurePlot(x=mtottrain[,-c(13)],y=mtottrain$classe,plot="strip")
```

## Preparatin for cross validation
Accuracy on a training set is often optimistic. To be able to do a better estimate,
I have chosen to split the original training set to a separate training and testing set.
70% of the data will be in this new training set and the rest in the testing set.

Estimating the test set accuray with the training set will hopefullt improve the accuarcy of the original testing set.

```{r, echo=FALSE}
inTrain <- createDataPartition(y=mtottrain$classe, p=0.7, list=FALSE)
mtrain <- mtottrain[inTrain,]
mtest <- mtottrain[-inTrain,]
dim(mtrain)
dim(mtest)
```

## Fitting different models to see which one gives the best accuracy.
I have chosen 3 of the methods which are used for categorization:
- Gradiant Boosting
- Random Foreset
- Support Vector Machines

```{r, echo=FALSE}
set.seed(55677)
modfit1 <- train(classe~., data=mtrain, method="gbm", verbose=FALSE)  
modfit2 <- train(classe~., data=mtrain, method="rf")  
library(e1071)
modfit3 <- svm(classe~., data=mtrain)
```

### Training result of Gradient Boosting model
```{r}
print (modfit1)
```

### Training result of Random Forest model
```{r}
print (modfit2)

```

##Making predictions for all 3 models on the training set

```{r}
pred1 <- predict(modfit1,mtest)
pred2 <- predict(modfit2,mtest)
pred3 <- predict(modfit3,mtest)

confusionMatrix(pred1,mtest$classe) 
confusionMatrix(pred2,mtest$classe) 
confusionMatrix(pred3,mtest$classe) 

```
### Prediction Summary for the training set
The Random Forest model gave the best prediction for the training set, the boosting model second and the svm model last.
Decided to use the Random Forest model do to predictions on the original testing set


### Making predictions for the original testing set
```{r}
# creating new dataframe with only the accelerometer data 
mtottest <- orgtest[,c(40,41,42,    #belt
                    63,64,65,    #arm
                    116,117,118, #dumbbell
                    154,155,156)] #forearm
                     
pred_final <- predict(modfit2,mtottest)

```


## Expected out of sample error
```{r}

table(pred2,mtest$classe)
sum(mtest$classe !=pred2)

sum(mtest$classe !=pred2) / length(mtest$classe)

```

